Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-16 ms.tex
Edited Section 3.2

We turn now to a discussion of our parameter inference analysis of projected galaxy clustering
in Decorated HOD models that include a treatment of galaxy assembly bias. In this work,
we consider only the simplest model of galaxy assembly bias, introducing only two new
parameters, $A_{\rm cen}$ and $A_{\rm sat}$, that describe the strength of central galaxy
and satellite galaxy assembly bias respectively. These parameters are limited to values
of $-1 \le A_{\rm cen,sat} \le 1$, and $A_{\rm cen,sat}=0$ when there is no galaxy assembly bias.
In this work, we use halo concentration as our secondary halo property, so $A_{\rm cen,sat}=1$
($A_{\rm cen,sat}=-1$) means that the mean number of galaxies per halo is maximally
correlated (anti-correlated) with halo concentration. The model and its implementation
in {\tt Halotools} is discussed further in Section~\ref{subsection:decorated} above and
in more detail in \citet{hearin_etal16}.

Examples of our fits are given in the right-hand panels of Figure~\ref{fig:Mr19samples},
Figure~\ref{fig:Mr20samples}, and Figure~\ref{fig:Mr21samples}. The general trend that
can be gleaned from these figures is that introducing assembly bias improves the ability
of the predicted two-point functions to match the measured two-point functions across the
transition from the one-halo (highly nonlinear) to two-halo (nearly linear) regimes near $\rp \sim 2\, h^{-1}{\mathrm{Mpc}}$.
This is most apparent for the $M_r < -20$ threshold sample shown in Fig.~\ref{fig:Mr20samples}.
Visually, these differences appear to be small; however, Table~\ref{table:parameters} shows that
they are statistically important.

The one-dimensional marginalized constraints on all parameters from these analyses
are given in the lowest row of each luminosity threshold grouping in Table~\ref{table:parameters}.
In cases where the posterior on a parameter is monotonic within the physical parameter
range, we quote an upper or lower limit on the parameter. Examples of one- and two-dimensional
visualizations of the posteriors from our analyses are shown in Figure~\ref{fig:Mr19ABtriangle},
Figure~\ref{fig:Mr20ABtriangle}, and Figure~\ref{fig:Mr21ABtriangle} for the
$M_r<-19$, $M_r<-20$, and $M_r<-21$ samples respectively.

@@

Table~\ref{table:parameters} and Figures~\ref{fig:Mr19ABtriangle}--\ref{fig:Mr21ABtriangle}
all make several simple, generic points. Introducing additional parameter
freedom associated with galaxy assembly bias generally increases the viable parameter
space, even for the subset of standard HOD parameters. Constraints on the
standard HOD parameters are generally less restrictive. This is exactly what is
expected from the introduction of additional parameter freedom.

@@

Beyond those generic conclusions, a few specific cases are worthy of further examination.
Consider the $M_r < -20$ sample. The inferred value of $A_{\rm cen} > 0.28$ at 99\%
confidence. In this case the data strongly prefer $A_{\rm cen} > 0$ and thus strongly prefer
galaxies to reside in halos of larger concentration at fixed halo mass. Figure~\ref{fig:directhodcomparison}
compares the inferred HODs for the $M_r<-20$ sample in a standard HOD analysis with the
inferred HOD from our decorated HOD analysis that includes parameters to describe
assembly bias. This is an explicit example of the degree to which the inferred relationship
between galaxies and halos can be altered by assembly bias.

Though the evidence for assembly bias is the most significant for the $M_r<-20$ sample,
there are hints of galaxy assembly bias in other samples. Satellite galaxies show a marginal preference
for occupying halos of higher concentration in the $M_r < -19$ threshold sample. The $M_r < -19.5$
sample exhibits weak preference for a positive correlation of galaxy occupation with halo concentration at fixed
mass for both satellite galaxies and central galaxies. Continuing upward with luminosity,
the $M_r < -20.5$ sample exhibits a significant preference for central galaxy assembly bias.
Lastly, there is no preference for either central galaxy or satellite galaxy assembly bias
for the $M_r < -21$ threshold sample, for which both $A_{\mathrm{cen}}$ and $A_{\mathrm{sat}}$
are consistent with zero well within 1$\sigma$.

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{table}
\begin{center}
{\renewcommand{\arraystretch}{1.3}
\renewcommand{\tabcolsep}{0.2cm}
\begin{tabular}{c c}
\hline
\hline
$M_r$ Threshold & $\Delta \mathrm{BIC}$ \\
\hline
-21 & -0.54 \\
-20.5 & 1.33 \\
-20 & 4.56\\
-19.5 & 0.26\\
-19 & 4.37\\
\hline
\end{tabular}
\medskip
\caption{
Change to the Bayesian Information Criterion, $\Delta \mathrm{BIC}$,
after introducing additional parametric freedom to accommodate galaxy
assembly bias. Sign convention is such that positive values favor models including
assembly bias, negative values favor standard HOD models with no assembly bias parameters. Changes in
the Bayesian Information Criterion $\vert \Delta \mathrm{BIC}\vert \ge 5$ strongly favor one model
over another.
}
 }
 \label{table:priors}
 \end{center}
\end{table}
%--------------------------------------------------------------------------------------------------------------------------------

@@

%------------------------------------------------
\bibliography{ms.bib}

>>>>

We turn now to a discussion of our parameter inference analysis of projected galaxy clustering
in decorated HOD models that include a treatment of galaxy assembly bias. In this work,
we consider only the simplest model of galaxy assembly bias, introducing only two new
parameters, $A_{\rm cen}$ and $A_{\rm sat}$, that describe the strength of central galaxy
and satellite galaxy assembly bias respectively. These parameters are limited to values
of $-1 \le A_{\rm cen,sat} \le 1$, and $A_{\rm cen,sat}=0$ when there is no galaxy assembly bias.
In this work, we use halo concentration as our secondary halo property, so $A_{\rm cen,sat}=1$
($A_{\rm cen,sat}=-1$) means that the mean number of galaxies per halo is maximally
correlated (anti-correlated) with halo concentration. The model and its implementation
in {\tt Halotools} \cite{halotools} is discussed further in Section~\ref{subsection:decorated} above 
and in more detail in \citet{hearin_etal16}.

Examples of our fits are given in the right-hand panels of Figure~\ref{fig:Mr19samples},
Figure~\ref{fig:Mr20samples}, and Figure~\ref{fig:Mr21samples}. The general trend that
can be gleaned from these figures is that introducing assembly bias improves the ability
of the predicted two-point functions to match the measured two-point functions across the
transition from the one-halo (highly nonlinear) to two-halo (nearly linear) regimes near 
$\rp \sim 2\, h^{-1}{\mathrm{Mpc}}$, an effect that could be anticipated by the known 
scale-dependence of galaxy assembly bias \citep{sunayama_etal16}. 
This is most apparent for the $M_r < -20$ threshold sample shown in Fig.~\ref{fig:Mr20samples}.
Visually, these differences appear to be small; however, Table~\ref{table:parameters} shows that
they are statistically important.

The one-dimensional marginalized constraints on all parameters from these analyses
are given in the lowest row of each luminosity threshold grouping in Table~\ref{table:parameters}.
In cases where the posterior distribution of either $A_{\rm cen}$ or $A_{\rm sat}$ is a monotonic within 
the physical parameter range, we quote an upper or lower limit on the parameter 
(e.g., see the posterior distribution for $A_{\rm cen}$ in Fig.~\ref{fig:Mr20ABtriangle}). 
Examples of one- and two-dimensional visualizations of the posteriors from our analyses are shown in 
Figure~\ref{fig:Mr19ABtriangle}, Figure~\ref{fig:Mr20ABtriangle}, and Figure~\ref{fig:Mr21ABtriangle} 
for the $M_r<-19$, $M_r<-20$, and $M_r<-21$ samples respectively.

@@

Table~\ref{table:parameters} and Figures~\ref{fig:Mr19ABtriangle}--\ref{fig:Mr21ABtriangle}
all make several simple, generic points. Introducing additional parameter
freedom associated with galaxy assembly bias generally increases the volume of the 
viable parameter space, even for the subset of five standard HOD parameters. 
Constraints on the standard HOD parameters are generally less restrictive in models that 
include assembly bias. This is exactly what is expected from the introduction of additional 
parameter freedom.

@@

Beyond those generic conclusions, a few specific cases are worthy of further examination.
Consider the $M_r < -20$ sample. The inferred value of $A_{\rm cen} > 0.28$ at 99\%
confidence. In this case the data strongly prefer $A_{\rm cen} > 0$ and thus strongly prefer
galaxies to reside in halos of larger concentration at fixed halo mass. Figure~\ref{fig:directhodcomparison}
compares the inferred HODs for the $M_r<-20$ sample in a standard HOD analysis with the
inferred HOD from our decorated HOD analysis that includes parameters to describe
assembly bias. This is an explicit example of the degree to which the inferred relationship
between galaxies and halos can be altered by assembly bias. In this case, the efficiency 
of forming a sufficiently luminous galaxy is a function that increases slowly over nearly two 
decades in halo mass in contrast to a standard analysis HOD where this efficiency rises 
rapidly over roughly one third of a decade in $\mvir$. Taking these results at face value, 
the reason that the HOD of central galaxies can vary slowly over several orders of magnitude 
in $\mvir$ is that larger galaxies form more efficiently in high-concentration halos at fixed 
mass.

Though the evidence for assembly bias is the most significant for the $M_r<-20$ sample,
there are hints of galaxy assembly bias in other samples. Satellite galaxies show a marginal preference
for occupying halos of higher concentration in the $M_r < -19$ threshold sample. The $M_r < -19.5$
sample exhibits weak preference for a positive correlation of galaxy occupation with halo concentration at fixed
mass for both satellite galaxies and central galaxies. Continuing upward with luminosity,
the $M_r < -20.5$ sample exhibits a significant preference for central galaxy assembly bias.
Lastly, there is no preference for either central galaxy or satellite galaxy assembly bias
for the $M_r < -21$ threshold sample, for which both $A_{\mathrm{cen}}$ and $A_{\mathrm{sat}}$
are consistent with zero well within 1$\sigma$. The implications of all of these results are too numerous 
to address in the present manuscript; however, we plan to address them in forthcoming publications.

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{table}
\begin{center}
{\renewcommand{\arraystretch}{1.3}
\renewcommand{\tabcolsep}{0.2cm}
\begin{tabular}{c c}
\hline
\hline
$M_r$ Threshold & $\Delta \mathrm{BIC}$ \\
\hline
-21 & -0.54 \\
-20.5 & 1.33 \\
-20 & 4.56\\
-19.5 & 0.26\\
-19 & 4.37\\
\hline
\end{tabular}
\medskip
\caption{
Change to the Bayesian Information Criterion, $\Delta \mathrm{BIC}$,
after introducing additional parametric freedom to accommodate galaxy
assembly bias. Sign convention is such that positive values favor models including
assembly bias, negative values favor standard HOD models with no assembly bias parameters. Changes in
the Bayesian Information Criterion $\vert \Delta \mathrm{BIC}\vert \ge 5$ strongly favor one model
over another.
}
 }
 \label{table:bic}
 \end{center}
\end{table}
%--------------------------------------------------------------------------------------------------------------------------------

@@

%------------------------------------------------
\bibliography{ms}
